---
title: "Elephant plots from raw processed data"
author: "K L Purves"
date: '`r format(Sys.time(), "%d %B, %Y")`'

output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: cerulean
code_folding: show

html_notebook:
  theme: cerulean
toc: yes
---

#### Clear global environment
```{r Clear global environment}
remove(list = ls())
```


# Description

This script will take the raw processed data from step one (i.e. a dataset that is well shaped and labelled, but not yet aggregated or with any appropriate issing data imputations etc) and create a series of plots to visualise the data. This will help for quick visual QCs and feasibility checks.

The code can be easily adapted to work on a clean dataset with missing values imputed. 

```{r Setup, libraries}

if(!require(ggplot2)){
  install.packages("ggplot2")
  library(ggplot2)
}

if(!require(dplyr)){
  install.packages("dplyr")
  library(dplyr)
}

if(!require(tidyr)){
  install.packages("tidyr")
  library(tidyr)
}

if(!require(cowplot)){
  install.packages("cowplot")
  library(cowplot)
}

if(!require(gtools)){
  install.packages("gtools")
  library(gtools)
}

```

# create plot ready data

separate out affective ratings and make a long dataframe, with summary data across the cohort.


## read in processed data from stage zero

```{r read in stage zero data and get the folder reference to save figures}

## file picker
filename_data <- file.choose()

## get the path to the directory where the raw data is stored
figure_path <- file.path(dirname(dirname(filename_data)),"Figures")

data_all<- read.csv(filename_data)

```


## create expectancy dataset

drop affective ratings, volume vars and CSpID

```{r drop affective}

expectancy <- data_all %>%
  select(!contains(c("volume", "CSpID","Affective")))


```

## create affective dataset
drop all expectancy ratings and volume vars
```{r drop expectancy}

affective <- data_all %>%
  select(contains(c("Subject_ID","Affective")))


```

## make plot dataset

Long format, with summary data (means across sample, 95% CIs)

### expectancy
make long, drop generalisation and trials that were not actually responded to due to drop out, make 999 (missed response) NA, create summary of data
```{r expectancy plot data}
expectancy_plot_data <- expectancy %>%
  pivot_longer(cols = !c(Subject_ID)) %>%
  separate(name, c("Phase","Stimulus","Trial"),convert = T) %>%
filter(complete.cases(.)) %>%
  filter(Phase != "Generalisation") %>%
  mutate(value = ifelse(value == 999,NA,value)) %>%
  group_by(Phase, Stimulus, Trial) %>%
  summarise(Mean = mean(value,na.rm = TRUE),
            SD = sd(value, na.rm = TRUE),
            N = n()) %>%
  mutate(SE = SD/sqrt(N),
         ci_low = Mean - qt(1-(0.05/2),N-1) * SE,
         ci_high = Mean + qt(1-(0.05/2),N-1) * SE) 
```

### affective
make long, drop g trials that were not actually responded to due to drop out, make 999 (missed response) NA, create summary of data
```{r affective plot data}

affective_plot_data <- affective %>%
  pivot_longer(cols = !c(Subject_ID)) %>%
  separate(name, c("Rating","Type","Stimulus","Phase"),convert = T) %>%
filter(complete.cases(.)) %>%
  mutate(value = ifelse(value == 999,NA,value)) %>%
  select(-Rating) %>%
  group_by(Type,Stimulus,Phase) %>%
  summarise(Mean = mean(value,na.rm = TRUE),
            SD = sd(value, na.rm = TRUE),
            N = n()) %>%
  mutate(SE = SD/sqrt(N),
         ci_low = Mean - qt(1-(0.05/2),N-1) * SE,
         ci_high = Mean + qt(1-(0.05/2),N-1) * SE) 

```

# Expectancy elephant plot

set up plot parameters such as palette and font sizes

```{r parameter specs}

PMpallette <- c("#002D64","#ACB9EA")  

### SET UP TEXT SIZE HERE ########

y.title <- 12
axis.text <-  10
strip.text <- 12

plot.title <- 16
plot.subtitle <- 12

# png resolution and units
res <- 700
units = 'in'

```

## plot

```{r elephant plot}


elephant <-  ggplot(expectancy_plot_data,
                     aes(Trial,Mean,
                         color=Stimulus))         +
  geom_point(size=2)                                        +
  geom_line(size =1)                                        +
  geom_ribbon(aes(ymin=ci_low, ymax=ci_high,
                  fill=Stimulus),
              linetype=2, alpha=0.1,
              size=.8)                                      +
  scale_fill_manual( values=c(PMpallette))                   +
  facet_grid(. ~ Phase, scales = "free_x",
             space = "free_x")              +
  theme_bw()                                                +
  scale_color_manual(values=PMpallette)                     +
  labs(title=paste("Preliminary Elephant plot"),
         y="Mean expectancy rating\n",
       subtitle = paste("dashed lines = 95% confidence intervals\nminimum n (all phases complete) = ", 
                        min(expectancy_plot_data$N),
                        "\nmaximum n (any data) = ",max(expectancy_plot_data$N) ))                                   +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.position = 'right',
        axis.title = element_text(face = "bold",
                                  size = y.title),
        panel.spacing.x=unit(0, "lines"))                     +
  
  scale_x_continuous(breaks = round(seq(min(expectancy_plot_data$Trial),
                                        max(expectancy_plot_data$Trial), by = 1),1))          +
  scale_y_continuous(breaks = seq(0,10,by=1))         +
  theme(plot.margin = unit(c(0.6, 0.6, 0.6, 0.6), "cm"))


elephant

```


# Affective rating plot

set up plot parameters such as palette and font sizes

## plot

```{r affective plot}


affective_figure <-  ggplot(affective_plot_data,
                     aes(Phase,Mean,
                         color=Stimulus))         +
  geom_point(size=2)                                        +
  geom_line(size =1)                                        +
  geom_ribbon(aes(ymin=ci_low, ymax=ci_high,
                  fill=Stimulus),
              linetype=2, alpha=0.1,
              size=.8)                                      +
  scale_fill_manual( values=c(PMpallette))                   +
  facet_grid(. ~ Type, scales = "free_x",
             space = "free_x")              +
  theme_bw()                                                +
  scale_color_manual(values=PMpallette)                     +
  labs(title=paste("ongoing data checks: affective ratings"),
         y="Mean affective rating\n",
       subtitle = paste("dashed lines = 95% confidence intervals\nminimum n (all phases complete) = ", 
                        min(affective_plot_data$N),
                        "\nmaximum n (any data) = ",max(affective_plot_data$N) ))                                   +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.position = 'right',
        axis.title = element_text(face = "bold",
                                  size = y.title),
        panel.spacing.x=unit(0, "lines"))                     +
  
  scale_x_continuous(breaks = round(seq(min(affective_plot_data$Phase),
                                        max(affective_plot_data$Phase), by = 1),1))          +
  scale_y_continuous(breaks = seq(0,10,by=1))         +
  theme(plot.margin = unit(c(0.6, 0.6, 0.6, 0.6), "cm"))


affective_figure

```


# save figures 

```{r save elephant}

# save figure in folder
png(paste0(figure_path,"/DataQC_ElephantPlot.png"),width=18,height=10,unit=units,res=res)
ggdraw(elephant)
dev.off()

# save figure in folder
png(paste0(figure_path,"/DataQC_AffectiveRatings.png"),width=18,height=10,unit=units,res=res)
ggdraw(affective_figure)
dev.off()

```

#### Clear global environment
```{r Clear global environment end}
remove(list = ls())
```
