---
title: "Elephant plots from raw processed data"
author: "K L Purves"
date: '`r format(Sys.time(), "%d %B, %Y")`'

output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: cerulean
code_folding: show

html_notebook:
  theme: cerulean
toc: yes
---

#### Clear global environment
```{r Clear global environment}
remove(list = ls())
```


# Description

This script will take the processed data from step two (i.e. a dataset that is well shaped and labelled, missing data imputed as necessary and aggregated) and create a series of plots to visualise the data. This will help for quick visual QCs and feasibility checks.

Elephant plots are then displayed per group.

Note, exclusions have not yet been applied.

The code can be easily adapted to work on a clean dataset with exclusions dropped, or to drop exclusions once they have been merged onto the dataset with correct IDs.



```{r Setup, libraries}

if(!require(ggplot2)){
  install.packages("ggplot2")
  library(ggplot2)
}

if(!require(dplyr)){
  install.packages("dplyr")
  library(dplyr)
}

if(!require(tidyr)){
  install.packages("tidyr")
  library(tidyr)
}

if(!require(cowplot)){
  install.packages("cowplot")
  library(cowplot)
}

if(!require(gtools)){
  install.packages("gtools")
  library(gtools)
}

```

# create plot ready data

separate out affective ratings and make a long dataframe, with summary data across the cohort.


## read in processed data

Read in the final dataset (called "FLAReApp_WithIDS_and_sumscores.csv" from the final_datasets folder. path should be datafiles/Final_datasets/)

```{r read in stage zero data and get the folder reference to save figures}

## file picker
filename_data <- file.choose()

## get the path to the directory where the raw data is stored
figure_path <- file.path(dirname(dirname(filename_data)),"Figures")

data_all<- read.csv(filename_data)

```


## create expectancy dataset

drop affective ratings, volume vars and CSpID, S4S variables, and mean summary variables.

Note, newgroup_4 = [1=trauma exposed control, 2=PTSD, 3=AUD, 4=both]; it uses a stricter cut off for alcohol use disorder than New2, so we drop new 2. Will retain PTSD/AUD columns to check these larger groups


```{r drop affective}

expectancy <- data_all %>%
  select(!contains(c("volume", "CSpID","Affective", "Mean","FirstThird","LastThird","Differential","Missing",
                     "to_consent_covid","timewanted_to_covid","from","tostatus","dna","yremoved",
                     "status_descr","tostatus","contact","tstatus",
                     "NewGroup_2")))


```

## create affective dataset
drop all expectancy ratings and volume vars
```{r drop expectancy}

affective <- data_all %>%
  rename(Subject_ID = "ID",
         AUD = "AUD4orMore") %>%
  select(contains(c("Subject_ID","Affective","NewGroup_4","PTSD","AUD")))



```

## make plot dataset

Long format, with summary data (means across sample, 95% CIs)

### expectancy plot data 
make long, drop generalisation and trials that were not actually responded to due to drop out, make 999 (missed response) NA, create summary of data

#### whole group
drop grouping variables
```{r expectancy plot data}
expectancy_plot_data <- expectancy %>%
  select(!NewGroup_4,!AUD4orMore,!PTSD) %>%
  pivot_longer(cols = !c(ID)) %>%
  separate(name, c("Phase","Stimulus","Trial"),convert = T) %>%
filter(complete.cases(.)) %>%
  filter(Phase != "Generalisation") %>%
  mutate(value = ifelse(value == 999,NA,value)) %>%
  group_by(Phase, Stimulus, Trial) %>%
  summarise(Mean = mean(value,na.rm = TRUE),
            SD = sd(value, na.rm = TRUE),
            N = n()) %>%
  mutate(SE = SD/sqrt(N),
         ci_low = Mean - qt(1-(0.05/2),N-1) * SE,
         ci_high = Mean + qt(1-(0.05/2),N-1) * SE) 
```

#### grouped by Trauma-exposed control, PTSD, AUD, PTSD+AUD

Group by NewGroup_4
```{r expectancy plot data new group}
expectancy_plot_data_NewGroup4 <- expectancy %>%
  select(!AUD4orMore,!PTSD) %>%
  pivot_longer(cols = !c(ID,NewGroup_4)) %>%
  separate(name, c("Phase","Stimulus","Trial"),convert = T) %>%
filter(complete.cases(.)) %>%
  filter(Phase != "Generalisation") %>%
  mutate(value = ifelse(value == 999,NA,value)) %>%
  group_by(NewGroup_4, Phase, Stimulus, Trial) %>%
  summarise(Mean = mean(value,na.rm = TRUE),
            SD = sd(value, na.rm = TRUE),
            N = n()) %>%
  mutate(SE = SD/sqrt(N),
         ci_low = Mean - qt(1-(0.05/2),N-1) * SE,
         ci_high = Mean + qt(1-(0.05/2),N-1) * SE) 

expectancy_plot_data_NewGroup4$NewGroup_4 <- factor(expectancy_plot_data_NewGroup4$NewGroup_4,
                                                    labels = c("Trauma-exposed control","PTSD","AUD","PTSD+AUD"))

names(expectancy_plot_data_NewGroup4)[1] <- "case_status"

# cut off confidence intervals at 1 and 10 for plotting. 

expectancy_plot_data_NewGroup4$ci_low <- ifelse(expectancy_plot_data_NewGroup4$ci_low < 1,1,expectancy_plot_data_NewGroup4$ci_low)
expectancy_plot_data_NewGroup4$ci_high <- ifelse(expectancy_plot_data_NewGroup4$ci_high > 9,9,expectancy_plot_data_NewGroup4$ci_high)
```

#### grouped by PTSD status

```{r expectancy plot data PTSD}
expectancy_plot_data_PTSD <- expectancy %>%
  select(!AUD4orMore,!NewGroup_4) %>%
  pivot_longer(cols = !c(ID,PTSD)) %>%
  separate(name, c("Phase","Stimulus","Trial"),convert = T) %>%
filter(complete.cases(.)) %>%
  filter(Phase != "Generalisation") %>%
  mutate(value = ifelse(value == 999,NA,value)) %>%
  group_by(PTSD, Phase, Stimulus, Trial) %>%
  summarise(Mean = mean(value,na.rm = TRUE),
            SD = sd(value, na.rm = TRUE),
            N = n()) %>%
  mutate(SE = SD/sqrt(N),
         ci_low = Mean - qt(1-(0.05/2),N-1) * SE,
         ci_high = Mean + qt(1-(0.05/2),N-1) * SE) 

expectancy_plot_data_PTSD$PTSD <- factor(expectancy_plot_data_PTSD$PTSD,
                                                    labels = c("Trauma-exposed control","PTSD"))

names(expectancy_plot_data_PTSD)[1] <- "case_status"
```

#### grouped by AUD status

```{r expectancy plot data AUD}

expectancy_plot_data_AUD <- expectancy %>%
  select(!PTSD,!NewGroup_4) %>%
  pivot_longer(cols = !c(ID,AUD4orMore)) %>%
  separate(name, c("Phase","Stimulus","Trial"),convert = T) %>%
filter(complete.cases(.)) %>%
  filter(Phase != "Generalisation") %>%
  mutate(value = ifelse(value == 999,NA,value)) %>%
  group_by(AUD4orMore, Phase, Stimulus, Trial) %>%
  summarise(Mean = mean(value,na.rm = TRUE),
            SD = sd(value, na.rm = TRUE),
            N = n()) %>%
  mutate(SE = SD/sqrt(N),
         ci_low = Mean - qt(1-(0.05/2),N-1) * SE,
         ci_high = Mean + qt(1-(0.05/2),N-1) * SE) 

expectancy_plot_data_AUD$AUD4orMore <- factor(expectancy_plot_data_AUD$AUD4orMore,
                                                    labels = c("Trauma-exposed control","AUD"))

names(expectancy_plot_data_AUD)[1] <- "case_status"
```

### affective: Whole group
make long, drop g trials that were not actually responded to due to drop out, make 999 (missed response) NA, create summary of data
```{r affective plot data}

affective_plot_data <- affective %>%
  select(!NewGroup_4,!PTSD,!AUD) %>%
  pivot_longer(cols = !c(Subject_ID)) %>%
  separate(name, c("Rating","Type","Stimulus","Phase"),convert = T) %>%
filter(complete.cases(.)) %>%
  mutate(value = ifelse(value == 999,NA,value)) %>%
  select(-Rating) %>%
  group_by(Type,Stimulus,Phase) %>%
  summarise(Mean = mean(value,na.rm = TRUE),
            SD = sd(value, na.rm = TRUE),
            N = n()) %>%
  mutate(SE = SD/sqrt(N),
         ci_low = Mean - qt(1-(0.05/2),N-1) * SE,
         ci_high = Mean + qt(1-(0.05/2),N-1) * SE) 

```

### affective: all four groups

```{r affective plot data four group}

affective_plot_data_NewGroup4 <- affective %>%
  select(!PTSD,!AUD) %>%
  pivot_longer(cols = !c(Subject_ID,NewGroup_4)) %>%
  separate(name, c("Rating","Type","Stimulus","Phase"),convert = T) %>%
filter(complete.cases(.)) %>%
  mutate(value = ifelse(value == 999,NA,value)) %>%
  select(-Rating) %>%
  group_by(NewGroup_4,Type,Stimulus,Phase) %>%
  summarise(Mean = mean(value,na.rm = TRUE),
            SD = sd(value, na.rm = TRUE),
            N = n()) %>%
  mutate(SE = SD/sqrt(N),
         ci_low = Mean - qt(1-(0.05/2),N-1) * SE,
         ci_high = Mean + qt(1-(0.05/2),N-1) * SE) 


affective_plot_data_NewGroup4$NewGroup_4<- factor(affective_plot_data_NewGroup4$NewGroup_4,
                                                    labels = c("Trauma-exposed control","PTSD","AUD","PTSD+AUD"))

names(affective_plot_data_NewGroup4)[1] <- "case_status"

# cut off confidence intervals at 1 and 10 for plotting. 

affective_plot_data_NewGroup4$ci_low <- ifelse(affective_plot_data_NewGroup4$ci_low < 1,1,affective_plot_data_NewGroup4$ci_low)
affective_plot_data_NewGroup4$ci_high <- ifelse(affective_plot_data_NewGroup4$ci_high > 9,9,affective_plot_data_NewGroup4$ci_high)
```

### affective: PTSD

```{r affective plot data PTSD}

affective_plot_data_PTSD <- affective %>%
  select(!NewGroup_4,!AUD) %>%
  pivot_longer(cols = !c(Subject_ID,PTSD)) %>%
  separate(name, c("Rating","Type","Stimulus","Phase"),convert = T) %>%
filter(complete.cases(.)) %>%
  mutate(value = ifelse(value == 999,NA,value)) %>%
  select(-Rating) %>%
  group_by(PTSD,Type,Stimulus,Phase) %>%
  summarise(Mean = mean(value,na.rm = TRUE),
            SD = sd(value, na.rm = TRUE),
            N = n()) %>%
  mutate(SE = SD/sqrt(N),
         ci_low = Mean - qt(1-(0.05/2),N-1) * SE,
         ci_high = Mean + qt(1-(0.05/2),N-1) * SE) 

affective_plot_data_PTSD$PTSD <- factor(affective_plot_data_PTSD$PTSD,
                                                    labels = c("Trauma-exposed control","PTSD"))

names(affective_plot_data_PTSD)[1] <- "case_status"

```

### affective: AUD

```{r affective plot data AUD}

affective_plot_data_AUD <- affective %>%
  select(!NewGroup_4,!PTSD) %>%
  pivot_longer(cols = !c(Subject_ID,AUD)) %>%
  separate(name, c("Rating","Type","Stimulus","Phase"),convert = T) %>%
filter(complete.cases(.)) %>%
  mutate(value = ifelse(value == 999,NA,value)) %>%
  select(-Rating) %>%
  group_by(AUD,Type,Stimulus,Phase) %>%
  summarise(Mean = mean(value,na.rm = TRUE),
            SD = sd(value, na.rm = TRUE),
            N = n()) %>%
  mutate(SE = SD/sqrt(N),
         ci_low = Mean - qt(1-(0.05/2),N-1) * SE,
         ci_high = Mean + qt(1-(0.05/2),N-1) * SE) 

affective_plot_data_AUD$AUD <- factor(affective_plot_data_AUD$AUD,
                                                    labels = c("Trauma-exposed control","AUD"))

names(affective_plot_data_AUD)[1] <- "case_status"
```

# Expectancy elephant plot

set up plot parameters such as palette and font sizes

```{r parameter specs}

PMpallette <- c("#002D64","#ACB9EA")  

### SET UP TEXT SIZE HERE ########

y.title <- 12
axis.text <-  10
strip.text <- 12

plot.title <- 16
plot.subtitle <- 12

# png resolution and units
res <- 700
units = 'in'

```

## plot whole group elephant

```{r elephant plot whole group}


elephant <-  ggplot(expectancy_plot_data,
                     aes(Trial,Mean,
                         color=Stimulus))         +
  geom_point(size=2)                                        +
  geom_line(size =1)                                        +
  geom_ribbon(aes(ymin=ci_low, ymax=ci_high,
                  fill=Stimulus),
              linetype=2, alpha=0.1,
              size=.8)                                      +
  scale_fill_manual( values=c(PMpallette))                   +
  facet_grid(. ~ Phase, scales = "free_x",
             space = "free_x")              +
  theme_bw()                                                +
  scale_color_manual(values=PMpallette)                     +
  labs(title=paste("Preliminary Elephant plot"),
         y="Mean expectancy rating\n",
       subtitle = paste("dashed lines = 95% confidence intervals\nminimum n (all phases complete) = ", 
                        min(expectancy_plot_data$N),
                        "\nmaximum n (any data) = ",max(expectancy_plot_data$N) ))                                   +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.position = 'right',
        axis.title = element_text(face = "bold",
                                  size = y.title),
        panel.spacing.x=unit(0, "lines"))                     +
  
  scale_x_continuous(breaks = round(seq(min(expectancy_plot_data$Trial),
                                        max(expectancy_plot_data$Trial), by = 1),1))          +
  scale_y_continuous(breaks = seq(0,10,by=1))         +
  theme(plot.margin = unit(c(0.6, 0.6, 0.6, 0.6), "cm"))


elephant

```

## plot all four groups

```{r elephant plot new group 4}

elephant_AllGroups <-  ggplot(expectancy_plot_data_NewGroup4,
                     aes(Trial,Mean,
                         color=Stimulus))         +
  geom_point(size=2)                                        +
  geom_line(size =1)                                        +
  geom_ribbon(aes(ymin=ci_low, ymax=ci_high,
                  fill=Stimulus),
              linetype=2, alpha=0.1,
              size=.8)                                      +
  scale_fill_manual( values=c(PMpallette))                   +
  facet_grid(case_status ~ Phase, scales = "free_x",
             space = "free_x")              +
  theme_bw()                                                +
  scale_color_manual(values=PMpallette)                     +
  labs(title=paste("Preliminary Elephant plot"),
       
         y="Mean expectancy rating\n",
       
       
       subtitle = paste("dashed lines = 95% confidence intervals\ntrauma-exposed controls n = ", 
                        max(expectancy_plot_data_NewGroup4$N[expectancy_plot_data_NewGroup4$case_status == "Trauma-exposed control"]),
                        "\nPTSD n = ",max(expectancy_plot_data_NewGroup4$N[expectancy_plot_data_NewGroup4$case_status == "PTSD"]),
                        "\nAUD n = ",max(expectancy_plot_data_NewGroup4$N[expectancy_plot_data_NewGroup4$case_status == "AUD"]),
                        "\nPTSD+AUD n = ",max(expectancy_plot_data_NewGroup4$N[expectancy_plot_data_NewGroup4$case_status == "PTSD+AUD"])
                        
                        ))                                   +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.position = 'right',
        axis.title = element_text(face = "bold",
                                  size = y.title),
        panel.spacing.x=unit(0, "lines"))                     +
  
  scale_x_continuous(breaks = round(seq(min(expectancy_plot_data$Trial),
                                        max(expectancy_plot_data$Trial), by = 1),1))          +
  scale_y_continuous(breaks = seq(0,10,by=1))         +
  theme(plot.margin = unit(c(0.6, 0.6, 0.6, 0.6), "cm"))


elephant_AllGroups

```
## plot PTSD vs trauma exposed controls

```{r elephant plot PTSD}

elephant_PTSD <-  ggplot(expectancy_plot_data_PTSD,
                     aes(Trial,Mean,
                         color=Stimulus))         +
  geom_point(size=2)                                        +
  geom_line(size =1)                                        +
  geom_ribbon(aes(ymin=ci_low, ymax=ci_high,
                  fill=Stimulus),
              linetype=2, alpha=0.1,
              size=.8)                                      +
  scale_fill_manual( values=c(PMpallette))                   +
  facet_grid(case_status ~ Phase, scales = "free_x",
             space = "free_x")              +
  theme_bw()                                                +
  scale_color_manual(values=PMpallette)                     +
  labs(title=paste("Preliminary Elephant plot"),
       
         y="Mean expectancy rating\n",
       
       
       subtitle = paste("dashed lines = 95% confidence intervals\ntrauma-exposed controls n = ", 
                        max(expectancy_plot_data_PTSD$N[expectancy_plot_data_PTSD$case_status == "Trauma-exposed control"]),
                        "\nPTSD n = ",max(expectancy_plot_data_PTSD$N[expectancy_plot_data_PTSD$case_status == "PTSD"])
                        
                        ))                                   +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.position = 'right',
        axis.title = element_text(face = "bold",
                                  size = y.title),
        panel.spacing.x=unit(0, "lines"))                     +
  
  scale_x_continuous(breaks = round(seq(min(expectancy_plot_data$Trial),
                                        max(expectancy_plot_data$Trial), by = 1),1))          +
  scale_y_continuous(breaks = seq(0,10,by=1))         +
  theme(plot.margin = unit(c(0.6, 0.6, 0.6, 0.6), "cm"))


elephant_PTSD

```

## plot AUD vs trauma exposed controls

```{r elephant plot AUD}

elephant_AUD <-  ggplot(expectancy_plot_data_AUD,
                     aes(Trial,Mean,
                         color=Stimulus))         +
  geom_point(size=2)                                        +
  geom_line(size =1)                                        +
  geom_ribbon(aes(ymin=ci_low, ymax=ci_high,
                  fill=Stimulus),
              linetype=2, alpha=0.1,
              size=.8)                                      +
  scale_fill_manual( values=c(PMpallette))                   +
  facet_grid(case_status ~ Phase, scales = "free_x",
             space = "free_x")              +
  theme_bw()                                                +
  scale_color_manual(values=PMpallette)                     +
  labs(title=paste("Preliminary Elephant plot"),
       
         y="Mean expectancy rating\n",
       
       
       subtitle = paste("dashed lines = 95% confidence intervals\ntrauma-exposed controls n = ", 
                        max(expectancy_plot_data_AUD$N[expectancy_plot_data_AUD$case_status == "Trauma-exposed control"]),
                        "\nAUD n = ",max(expectancy_plot_data_AUD$N[expectancy_plot_data_AUD$case_status == "AUD"])
                        
                        ))                                   +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.position = 'right',
        axis.title = element_text(face = "bold",
                                  size = y.title),
        panel.spacing.x=unit(0, "lines"))                     +
  
  scale_x_continuous(breaks = round(seq(min(expectancy_plot_data$Trial),
                                        max(expectancy_plot_data$Trial), by = 1),1))          +
  scale_y_continuous(breaks = seq(0,10,by=1))         +
  theme(plot.margin = unit(c(0.6, 0.6, 0.6, 0.6), "cm"))


elephant_AUD

```

# Affective rating plot

set up plot parameters such as palette and font sizes

## plot: Whole group 

```{r affective plot whole sample}


affective_figure <-  ggplot(affective_plot_data,
                     aes(Phase,Mean,
                         color=Stimulus))         +
  geom_point(size=2)                                        +
  geom_line(size =1)                                        +
  geom_ribbon(aes(ymin=ci_low, ymax=ci_high,
                  fill=Stimulus),
              linetype=2, alpha=0.1,
              size=.8)                                      +
  scale_fill_manual( values=c(PMpallette))                   +
  facet_grid(. ~ Type, scales = "free_x",
             space = "free_x")              +
  theme_bw()                                                +
  scale_color_manual(values=PMpallette)                     +
  labs(title=paste("ongoing data checks: affective ratings"),
         y="Mean affective rating\n",
       subtitle = paste("dashed lines = 95% confidence intervals\nminimum n (all phases complete) = ", 
                        min(affective_plot_data$N),
                        "\nmaximum n (any data) = ",max(affective_plot_data$N) ))                                   +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.position = 'right',
        axis.title = element_text(face = "bold",
                                  size = y.title),
        panel.spacing.x=unit(0, "lines"))                     +
  
  scale_x_continuous(breaks = round(seq(min(affective_plot_data$Phase),
                                        max(affective_plot_data$Phase), by = 1),1))          +
  scale_y_continuous(breaks = seq(0,10,by=1))         +
  theme(plot.margin = unit(c(0.6, 0.6, 0.6, 0.6), "cm"))


affective_figure

```

## plot: New Group 4

```{r affective plot 4 group}


affective_NewGroup4 <-  ggplot(affective_plot_data_NewGroup4,
                     aes(Phase,Mean,
                         color=Stimulus))         +
  geom_point(size=2)                                        +
  geom_line(size =1)                                        +
  geom_ribbon(aes(ymin=ci_low, ymax=ci_high,
                  fill=Stimulus),
              linetype=2, alpha=0.1,
              size=.8)                                      +
  scale_fill_manual( values=c(PMpallette))                   +
  facet_grid(case_status ~ Type, scales = "free_x",
             space = "free_x")              +
  theme_bw()                                                +
  scale_color_manual(values=PMpallette)                     +
  labs(title=paste("ongoing data checks: affective ratings"),
         y="Mean affective rating\n",
       
       subtitle = paste("dashed lines = 95% confidence intervals\ntrauma-exposed controls n = ", 
                        max(affective_plot_data_NewGroup4$N[affective_plot_data_NewGroup4$case_status == "Trauma-exposed control"]),
                        "\nPTSD n = ",max(affective_plot_data_NewGroup4$N[affective_plot_data_NewGroup4$case_status == "PTSD"]),
                        "\nAUD n = ",max(affective_plot_data_NewGroup4$N[affective_plot_data_NewGroup4$case_status == "AUD"]),
                        "\nPTSD+AUD n = ",max(affective_plot_data_NewGroup4$N[affective_plot_data_NewGroup4$case_status == "PTSD+AUD"])
                        
                        ))  +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.position = 'right',
        axis.title = element_text(face = "bold",
                                  size = y.title),
        panel.spacing.x=unit(0, "lines"))                     +
  
  scale_x_continuous(breaks = round(seq(min(affective_plot_data$Phase),
                                        max(affective_plot_data$Phase), by = 1),1))          +
  scale_y_continuous(breaks = seq(0,10,by=1))         +
  theme(plot.margin = unit(c(0.6, 0.6, 0.6, 0.6), "cm"))


affective_NewGroup4

```

## plot: PTSD

```{r affective plot PTSD}


affective_PTSD <-  ggplot(affective_plot_data_PTSD,
                     aes(Phase,Mean,
                         color=Stimulus))         +
  geom_point(size=2)                                        +
  geom_line(size =1)                                        +
  geom_ribbon(aes(ymin=ci_low, ymax=ci_high,
                  fill=Stimulus),
              linetype=2, alpha=0.1,
              size=.8)                                      +
  scale_fill_manual( values=c(PMpallette))                   +
  facet_grid(case_status ~ Type, scales = "free_x",
             space = "free_x")              +
  theme_bw()                                                +
  scale_color_manual(values=PMpallette)                     +
  labs(title=paste("ongoing data checks: affective ratings"),
         y="Mean affective rating\n",
       
       subtitle = paste("dashed lines = 95% confidence intervals\ntrauma-exposed controls n = ", 
                        max(affective_plot_data_PTSD$N[affective_plot_data_PTSD$case_status == "Trauma-exposed control"]),
                        "\nPTSD n = ",max(affective_plot_data_PTSD$N[affective_plot_data_PTSD$case_status == "PTSD"])
                        
                        ))  +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.position = 'right',
        axis.title = element_text(face = "bold",
                                  size = y.title),
        panel.spacing.x=unit(0, "lines"))                     +
  
  scale_x_continuous(breaks = round(seq(min(affective_plot_data$Phase),
                                        max(affective_plot_data$Phase), by = 1),1))          +
  scale_y_continuous(breaks = seq(0,10,by=1))         +
  theme(plot.margin = unit(c(0.6, 0.6, 0.6, 0.6), "cm"))


affective_PTSD

```
## plot: AUD

```{r affective plot AUD}


affective_AUD <-  ggplot(affective_plot_data_AUD,
                     aes(Phase,Mean,
                         color=Stimulus))         +
  geom_point(size=2)                                        +
  geom_line(size =1)                                        +
  geom_ribbon(aes(ymin=ci_low, ymax=ci_high,
                  fill=Stimulus),
              linetype=2, alpha=0.1,
              size=.8)                                      +
  scale_fill_manual( values=c(PMpallette))                   +
  facet_grid(case_status ~ Type, scales = "free_x",
             space = "free_x")              +
  theme_bw()                                                +
  scale_color_manual(values=PMpallette)                     +
  labs(title=paste("ongoing data checks: affective ratings"),
         y="Mean affective rating\n",
       
       subtitle = paste("dashed lines = 95% confidence intervals\ntrauma-exposed controls n = ", 
                        max(affective_plot_data_AUD$N[affective_plot_data_AUD$case_status == "Trauma-exposed control"]),
                        "\nAUD n = ",max(affective_plot_data_AUD$N[affective_plot_data_AUD$case_status == "AUD"])
                        
                        ))  +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.position = 'right',
        axis.title = element_text(face = "bold",
                                  size = y.title),
        panel.spacing.x=unit(0, "lines"))                     +
  
  scale_x_continuous(breaks = round(seq(min(affective_plot_data$Phase),
                                        max(affective_plot_data$Phase), by = 1),1))          +
  scale_y_continuous(breaks = seq(0,10,by=1))         +
  theme(plot.margin = unit(c(0.6, 0.6, 0.6, 0.6), "cm"))


affective_AUD

```

# save figures 

```{r save elephant}

# save expectancy figures in folder
png(paste0(figure_path,"/DataQC_ElephantPlot_WholeSample.png"),width=18,height=10,unit=units,res=res)
ggdraw(elephant)
dev.off()

png(paste0(figure_path,"/DataQC_ElephantPlot_NewGroup4.png"),width=18,height=10,unit=units,res=res)
ggdraw(elephant_AllGroups)
dev.off()

png(paste0(figure_path,"/DataQC_ElephantPlot_PTSD.png"),width=18,height=10,unit=units,res=res)
ggdraw(elephant_PTSD)
dev.off()

png(paste0(figure_path,"/DataQC_ElephantPlot_AUD.png"),width=18,height=10,unit=units,res=res)
ggdraw(elephant_AUD)
dev.off()

# save affective figures in folder
png(paste0(figure_path,"/DataQC_AffectiveRatings_WholeSample.png"),width=18,height=10,unit=units,res=res)
ggdraw(affective_figure)
dev.off()

png(paste0(figure_path,"/DataQC_AffectiveRatings_NewGroup4.png"),width=18,height=10,unit=units,res=res)
ggdraw(affective_NewGroup4)
dev.off()

png(paste0(figure_path,"/DataQC_AffectiveRatings_PTSD.png"),width=18,height=10,unit=units,res=res)
ggdraw(affective_PTSD)
dev.off()

png(paste0(figure_path,"/DataQC_AffectiveRatings_AUD.png"),width=18,height=10,unit=units,res=res)
ggdraw(affective_AUD)
dev.off()

```

#### Clear global environment
```{r Clear global environment end}
remove(list = ls())
```
