---
title: "Preliminary group comparisons of FLARe app data"
author: "K L Purves"
date: '`r format(Sys.time(), "%d %B, %Y")`'

output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: cerulean
code_folding: show

html_notebook:
  theme: cerulean
toc: yes
---

# Description

Preliminary group comparisons investigating group differences between different data subgroups in:

1. Whole phase means    
2. First third of each phase (Acquisition and Extinction)   
3. Last third means (Acquisition and Extinction)    
4. Differentials for all of the above.

# Setup

#### Clear global environment
```{r Clear global environment}
remove(list = ls())
```

```{r Setup, libraries}

if(!require(gtsummary)){
  install.packages("gtsummary")
  library(gtsummary)
}

if(!require(ggplot2)){
  install.packages("ggplot2")
  library(ggplot2)
}

if(!require(psych)){
  install.packages("psych")
  library(gtools)
}

if(!require(dplyr)){
  install.packages("dplyr")
  library(dplyr)
}

if(!require(tidyr)){
  install.packages("tidyr")
  library(tidyr)
}

```

## read in processed data

Read in the final dataset (called "FLAReApp_WithIDS_and_sumscores.csv" from the final_datasets folder. path should be datafiles/Final_datasets/)

```{r read in stage zero data and get the folder reference to save figures}

## file picker
filename_data <- file.choose()

## get the path to the directory where the raw data is stored
figure_path <- file.path(dirname(dirname(filename_data)),"Figures")
table_path <- file.path(dirname(dirname(filename_data)),"Tables")

data_all<- read.csv(filename_data)


data_all$NewGroup_4 <- factor(data_all$NewGroup_4,
                              labels = c("Trauma-exposed control","PTSD","AUD","PTSD+AUD"))
```

# Tables showing means and standard deviations

## New Group 4

```{r}
means_4group <- data_all %>%
  select(NewGroup_4,Acquisition_CSPlus_Mean,Acquisition_CSMinus_Mean,
         Extinction_CSPlus_Mean,Extinction_CSMinus_Mean,
         ReturnOfFear_CSPlus_Mean,ReturnOfFear_CSMinus_Mean,) %>%
  tbl_summary(NewGroup_4,
              statistic = list(all_continuous() ~ "{mean} ({sd})")) 
means_4group 

means_4group  %>%
  as_gt() %>% 
  gt::gtsave(
    filename = file.path(table_path,"WholePhaseMeans_NewGroup4.png")
  )
```


```{r}
first_4group <- data_all %>%
  select(NewGroup_4,Acquisition_CSPlus_FirstThird,Acquisition_CSMinus_FirstThird,
         Extinction_CSPlus_FirstThird,Extinction_CSMinus_FirstThird) %>%
  tbl_summary(NewGroup_4,
              statistic = list(all_continuous() ~ "{mean} ({sd})")) 
first_4group 

first_4group  %>%
  as_gt() %>% 
  gt::gtsave(
    filename = file.path(table_path,"FirstThirdMeans_NewGroup4.png")
  )
```

```{r}
last_4group <- data_all %>%
  select(NewGroup_4,Acquisition_CSPlus_LastThird,Acquisition_CSMinus_LastThird,
         Extinction_CSPlus_LastThird,Extinction_CSMinus_LastThird) %>%
  tbl_summary(NewGroup_4,
              statistic = list(all_continuous() ~ "{mean} ({sd})")) 
last_4group 

last_4group  %>%
  as_gt() %>% 
  gt::gtsave(
    filename = file.path(table_path,"LastThirdMeans_NewGroup4.png")
  )
```
## PTSD

```{r}
means_PTSD<- data_all %>%
  select(PTSD,Acquisition_CSPlus_Mean,Acquisition_CSMinus_Mean,
         Extinction_CSPlus_Mean,Extinction_CSMinus_Mean,
         ReturnOfFear_CSPlus_Mean,ReturnOfFear_CSMinus_Mean,) %>%
  tbl_summary(PTSD,
              statistic = list(all_continuous() ~ "{mean} ({sd})")) 
means_PTSD 

means_PTSD  %>%
  as_gt() %>% 
  gt::gtsave(
    filename = file.path(table_path,"WholePhaseMeans_PTSD.png")
  )
```

```{r}
first_PTSD<- data_all %>%
  select(PTSD,Acquisition_CSPlus_FirstThird,Acquisition_CSMinus_FirstThird,
         Extinction_CSPlus_FirstThird,Extinction_CSMinus_FirstThird) %>%
  tbl_summary(PTSD,
              statistic = list(all_continuous() ~ "{mean} ({sd})")) 
first_PTSD 

first_PTSD  %>%
  as_gt() %>% 
  gt::gtsave(
    filename = file.path(table_path,"FirstThirdMeans_PTSD.png")
  )
```

```{r}
last_PTSD<- data_all %>%
  select(PTSD,Acquisition_CSPlus_LastThird,Acquisition_CSMinus_LastThird,
         Extinction_CSPlus_LastThird,Extinction_CSMinus_LastThird) %>%
  tbl_summary(PTSD,
              statistic = list(all_continuous() ~ "{mean} ({sd})")) 
last_PTSD 

last_PTSD  %>%
  as_gt() %>% 
  gt::gtsave(
    filename = file.path(table_path,"LastThirdMeans_PTSD.png")
  )
```

## AUD

```{r}
means_AUD<- data_all %>%
  select(AUD4orMore,Acquisition_CSPlus_Mean,Acquisition_CSMinus_Mean,
         Extinction_CSPlus_Mean,Extinction_CSMinus_Mean,
         ReturnOfFear_CSPlus_Mean,ReturnOfFear_CSMinus_Mean,) %>%
  tbl_summary(AUD4orMore,
              statistic = list(all_continuous() ~ "{mean} ({sd})")) 
means_AUD 

means_AUD  %>%
  as_gt() %>% 
  gt::gtsave(
    filename = file.path(table_path,"WholePhaseMeans_AUD.png")
  )
```

```{r}
first_AUD<- data_all %>%
  select(AUD4orMore,Acquisition_CSPlus_FirstThird,Acquisition_CSMinus_FirstThird,
         Extinction_CSPlus_FirstThird,Extinction_CSMinus_FirstThird) %>%
  tbl_summary(AUD4orMore,
              statistic = list(all_continuous() ~ "{mean} ({sd})")) 
first_AUD 

first_AUD  %>%
  as_gt() %>% 
  gt::gtsave(
    filename = file.path(table_path,"FirstThirdMeans_AUD.png")
  )
```

```{r}
last_AUD<- data_all %>%
  select(AUD4orMore,Acquisition_CSPlus_LastThird,Acquisition_CSMinus_LastThird,
         Extinction_CSPlus_LastThird,Extinction_CSMinus_LastThird) %>%
  tbl_summary(AUD4orMore,
              statistic = list(all_continuous() ~ "{mean} ({sd})")) 
last_AUD 

last_AUD  %>%
  as_gt() %>% 
  gt::gtsave(
    filename = file.path(table_path,"LastThirdMeans_AUD.png")
  )
```

# Expectancy means bar plots {.tabset}

## set up common plot features

```{r parameter specs}

PMpallette <- c("#002D64","#ACB9EA")  

### SET UP TEXT SIZE HERE ########

y.title <- 12
axis.text <-  10
strip.text <- 12

plot.title <- 16
plot.subtitle <- 12

# png resolution and units
res <- 700
units = 'in'

dodge <- position_dodge(width = 0.85)

```


## New Group 4

### Create tidy plotting dataset 

This chunk creates a dataset with group means, SE, confidence intervals for each stimulus, phase and summary score type (overall mean, first third, last third)

```{r}

expectancy_plot_data <- data_all %>%
  select(ID,NewGroup_4,Acquisition_CSMinus_Mean,Acquisition_CSPlus_Mean,
         Extinction_CSMinus_Mean, Extinction_CSPlus_Mean,
         ReturnOfFear_CSMinus_Mean, ReturnOfFear_CSPlus_Mean,
         Acquisition_CSMinus_FirstThird,Acquisition_CSPlus_FirstThird,
         Extinction_CSMinus_FirstThird,Extinction_CSPlus_FirstThird,
         Acquisition_CSMinus_LastThird,Acquisition_CSPlus_LastThird,
         Extinction_CSMinus_LastThird,Extinction_CSPlus_LastThird) %>%
  pivot_longer(cols = !c(ID,NewGroup_4)) %>%
  separate(name, c("Phase","Stimulus","Variable"),convert = T) %>%
  filter(complete.cases(.)) %>%
  group_by(NewGroup_4, Phase, Stimulus, Variable) %>%
  summarise(Mean = mean(value,na.rm = TRUE),
            SD = sd(value, na.rm = TRUE),
            N = n()) %>%
  mutate(SE = SD/sqrt(N),
         ci_low = Mean - qt(1-(0.05/2),N-1) * SE,
         ci_high = Mean + qt(1-(0.05/2),N-1) * SE) 


expectancy_plot_data$NewGroup_4 <- factor(expectancy_plot_data$NewGroup_4,
                                                    labels = c("Trauma-exposed control","PTSD","AUD","PTSD+AUD"))

names(expectancy_plot_data)[1] <- "case_status"


expectancy_plot_data 

```

### bar plots: whole phase mean

```{r}
newGroup_bar <-  ggplot(expectancy_plot_data %>%
                          filter(Variable == "Mean"),
                     aes(case_status, Mean,
                         fill = Stimulus))  +
  geom_bar(position = "dodge",
           width = 0.9,
           stat="identity") +
  geom_errorbar(aes(ymin =  Mean - SE,
                    ymax = Mean + SE),
                position = position_dodge(width=.9),
                width = .1)                          +
  facet_wrap(. ~ Phase,
             nrow =2)                              +
  theme_bw()                                         +
  scale_fill_manual(values=PMpallette)                +        
  labs(title=paste("Mean expectancy rating across the whole phase"),
       
         y="Mean expectancy rating\n",
       
       
       subtitle = paste("Bars indicate standard error of the mean\n\ntrauma-exposed controls n = ", 
                        max(expectancy_plot_data$N[expectancy_plot_data$case_status == "Trauma-exposed control"]),
                        "\nPTSD n = ",max(expectancy_plot_data$N[expectancy_plot_data$case_status == "PTSD"]),
                        "\nAUD n = ",max(expectancy_plot_data$N[expectancy_plot_data$case_status == "AUD"]),
                        "\nPTSD+AUD n = ",max(expectancy_plot_data$N[expectancy_plot_data$case_status == "PTSD+AUD"])
                        
                        ))                             +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.key = element_rect(size = 3),
        legend.key.size = unit(3, 'lines'),
        axis.title.y =  element_text(face = "bold",
                                     size = y.title),
        axis.title.x =  element_blank(),
        panel.spacing.x=unit(0, "lines"))     + 
  scale_y_continuous(limits = c(0,9),
                     breaks = c(seq(0,9)))    

newGroup_bar

```

### bar plots: first third

```{r}
newGroup_bar_first <-  ggplot(expectancy_plot_data %>%
                          filter(Variable == "FirstThird"),
                     aes(case_status, Mean,
                         fill = Stimulus))  +
  geom_bar(position = "dodge",
           width = 0.9,
           stat="identity") +
  geom_errorbar(aes(ymin =  Mean - SE,
                    ymax = Mean + SE),
                position = position_dodge(width=.9),
                width = .1)                          +
  facet_wrap(. ~ Phase,
             nrow =2)                              +
  theme_bw()                                         +
  scale_fill_manual(values=PMpallette)                +        
  labs(title=paste("Mean expectancy rating across the First third of the phase"),
       
         y="Mean expectancy rating\n",
       
       
       subtitle = paste("Bars indicate standard error of the mean\n\ntrauma-exposed controls n = ", 
                        max(expectancy_plot_data$N[expectancy_plot_data$case_status == "Trauma-exposed control"]),
                        "\nPTSD n = ",max(expectancy_plot_data$N[expectancy_plot_data$case_status == "PTSD"]),
                        "\nAUD n = ",max(expectancy_plot_data$N[expectancy_plot_data$case_status == "AUD"]),
                        "\nPTSD+AUD n = ",max(expectancy_plot_data$N[expectancy_plot_data$case_status == "PTSD+AUD"])
                        
                        ))                             +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.key = element_rect(size = 3),
        legend.key.size = unit(3, 'lines'),
        axis.title.y =  element_text(face = "bold",
                                     size = y.title),
        axis.title.x =  element_blank(),
        panel.spacing.x=unit(0, "lines"))     + 
  scale_y_continuous(limits = c(0,9),
                     breaks = c(seq(0,9)))    

newGroup_bar_first

```
### bar plots: final third

```{r}
newGroup_bar_final <-  ggplot(expectancy_plot_data %>%
                          filter(Variable == "LastThird"),
                     aes(case_status, Mean,
                         fill = Stimulus))  +
  geom_bar(position = "dodge",
           width = 0.9,
           stat="identity") +
  geom_errorbar(aes(ymin =  Mean - SE,
                    ymax = Mean + SE),
                position = position_dodge(width=.9),
                width = .1)                          +
  facet_wrap(. ~ Phase,
             nrow =2)                              +
  theme_bw()                                         +
  scale_fill_manual(values=PMpallette)                +        
  labs(title=paste("Mean expectancy rating across the Last third of the phase"),
       
         y="Mean expectancy rating\n",
       
       
       subtitle = paste("Bars indicate standard error of the mean\n\ntrauma-exposed controls n = ", 
                        max(expectancy_plot_data$N[expectancy_plot_data$case_status == "Trauma-exposed control"]),
                        "\nPTSD n = ",max(expectancy_plot_data$N[expectancy_plot_data$case_status == "PTSD"]),
                        "\nAUD n = ",max(expectancy_plot_data$N[expectancy_plot_data$case_status == "AUD"]),
                        "\nPTSD+AUD n = ",max(expectancy_plot_data$N[expectancy_plot_data$case_status == "PTSD+AUD"])
                        
                        ))                             +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.key = element_rect(size = 3),
        legend.key.size = unit(3, 'lines'),
        axis.title.y =  element_text(face = "bold",
                                     size = y.title),
        axis.title.x =  element_blank(),
        panel.spacing.x=unit(0, "lines"))     + 
  scale_y_continuous(limits = c(0,9),
                     breaks = c(seq(0,9)))    

newGroup_bar_final

```

### PTSD

### Create tidy plotting dataset 

This chunk creates a dataset with group means, SE, confidence intervals for each stimulus, phase and summary score type (overall mean, first third, last third)

```{r}

expectancy_plot_data <- data_all %>%
  select(ID,PTSD,Acquisition_CSMinus_Mean,Acquisition_CSPlus_Mean,
         Extinction_CSMinus_Mean, Extinction_CSPlus_Mean,
         ReturnOfFear_CSMinus_Mean, ReturnOfFear_CSPlus_Mean,
         Acquisition_CSMinus_FirstThird,Acquisition_CSPlus_FirstThird,
         Extinction_CSMinus_FirstThird,Extinction_CSPlus_FirstThird,
         Acquisition_CSMinus_LastThird,Acquisition_CSPlus_LastThird,
         Extinction_CSMinus_LastThird,Extinction_CSPlus_LastThird) %>%
  pivot_longer(cols = !c(ID,PTSD)) %>%
  separate(name, c("Phase","Stimulus","Variable"),convert = T) %>%
  filter(complete.cases(.)) %>%
  group_by(PTSD, Phase, Stimulus, Variable) %>%
  summarise(Mean = mean(value,na.rm = TRUE),
            SD = sd(value, na.rm = TRUE),
            N = n()) %>%
  mutate(SE = SD/sqrt(N),
         ci_low = Mean - qt(1-(0.05/2),N-1) * SE,
         ci_high = Mean + qt(1-(0.05/2),N-1) * SE) 


expectancy_plot_data$PTSD <- factor(expectancy_plot_data$PTSD,
                                                    labels = c("Trauma-exposed control","PTSD"))

names(expectancy_plot_data)[1] <- "case_status"


expectancy_plot_data 
```
### bar plots: whole phase mean

```{r}
PTSD_bar <-  ggplot(expectancy_plot_data %>%
                          filter(Variable == "Mean"),
                     aes(case_status, Mean,
                         fill = Stimulus))  +
  geom_bar(position = "dodge",
           width = 0.9,
           stat="identity") +
  geom_errorbar(aes(ymin =  Mean - SE,
                    ymax = Mean + SE),
                position = position_dodge(width=.9),
                width = .1)                          +
  facet_wrap(. ~ Phase,
             nrow =2)                              +
  theme_bw()                                         +
  scale_fill_manual(values=PMpallette)                +        
  labs(title=paste("Mean expectancy rating across the whole phase"),
       
         y="Mean expectancy rating\n",
       
       
       subtitle = paste("Bars indicate standard error of the mean\n\ntrauma-exposed controls n = ", 
                        max(expectancy_plot_data$N[expectancy_plot_data$case_status == "Trauma-exposed control"]),
                        "\nPTSD n = ",max(expectancy_plot_data$N[expectancy_plot_data$case_status == "PTSD"])
                        
                        ))                             +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.key = element_rect(size = 3),
        legend.key.size = unit(3, 'lines'),
        axis.title.y =  element_text(face = "bold",
                                     size = y.title),
        axis.title.x =  element_blank(),
        panel.spacing.x=unit(0, "lines"))     + 
  scale_y_continuous(limits = c(0,9),
                     breaks = c(seq(0,9)))    

PTSD_bar

```

### bar plots: first third

```{r}
PTSD_bar_first <-  ggplot(expectancy_plot_data %>%
                          filter(Variable == "FirstThird"),
                     aes(case_status, Mean,
                         fill = Stimulus))  +
  geom_bar(position = "dodge",
           width = 0.9,
           stat="identity") +
  geom_errorbar(aes(ymin =  Mean - SE,
                    ymax = Mean + SE),
                position = position_dodge(width=.9),
                width = .1)                          +
  facet_wrap(. ~ Phase,
             nrow =2)                              +
  theme_bw()                                         +
  scale_fill_manual(values=PMpallette)                +        
  labs(title=paste("Mean expectancy rating across the First third of the phase"),
       
         y="Mean expectancy rating\n",
       
       
       subtitle = paste("Bars indicate standard error of the mean\n\ntrauma-exposed controls n = ", 
                        max(expectancy_plot_data$N[expectancy_plot_data$case_status == "Trauma-exposed control"]),
                        "\nPTSD n = ",max(expectancy_plot_data$N[expectancy_plot_data$case_status == "PTSD"])
                        
                        ))                             +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.key = element_rect(size = 3),
        legend.key.size = unit(3, 'lines'),
        axis.title.y =  element_text(face = "bold",
                                     size = y.title),
        axis.title.x =  element_blank(),
        panel.spacing.x=unit(0, "lines"))     + 
  scale_y_continuous(limits = c(0,9),
                     breaks = c(seq(0,9)))    

PTSD_bar_first

```
### bar plots: final third

```{r}
PTSD_bar_final <-  ggplot(expectancy_plot_data %>%
                          filter(Variable == "LastThird"),
                     aes(case_status, Mean,
                         fill = Stimulus))  +
  geom_bar(position = "dodge",
           width = 0.9,
           stat="identity") +
  geom_errorbar(aes(ymin =  Mean - SE,
                    ymax = Mean + SE),
                position = position_dodge(width=.9),
                width = .1)                          +
  facet_wrap(. ~ Phase,
             nrow =2)                              +
  theme_bw()                                         +
  scale_fill_manual(values=PMpallette)                +        
  labs(title=paste("Mean expectancy rating across the Last third of the phase"),
       
         y="Mean expectancy rating\n",
       
       
       subtitle = paste("Bars indicate standard error of the mean\n\ntrauma-exposed controls n = ", 
                        max(expectancy_plot_data$N[expectancy_plot_data$case_status == "Trauma-exposed control"]),
                        "\nPTSD n = ",max(expectancy_plot_data$N[expectancy_plot_data$case_status == "PTSD"])
                        ))                             +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.key = element_rect(size = 3),
        legend.key.size = unit(3, 'lines'),
        axis.title.y =  element_text(face = "bold",
                                     size = y.title),
        axis.title.x =  element_blank(),
        panel.spacing.x=unit(0, "lines"))     + 
  scale_y_continuous(limits = c(0,9),
                     breaks = c(seq(0,9)))    

PTSD_bar_final

```

### AUD

### Create tidy plotting dataset 

This chunk creates a dataset with group means, SE, confidence intervals for each stimulus, phase and summary score type (overall mean, first third, last third)

```{r}

expectancy_plot_data <- data_all %>%
  select(ID,AUD4orMore,Acquisition_CSMinus_Mean,Acquisition_CSPlus_Mean,
         Extinction_CSMinus_Mean, Extinction_CSPlus_Mean,
         ReturnOfFear_CSMinus_Mean, ReturnOfFear_CSPlus_Mean,
         Acquisition_CSMinus_FirstThird,Acquisition_CSPlus_FirstThird,
         Extinction_CSMinus_FirstThird,Extinction_CSPlus_FirstThird,
         Acquisition_CSMinus_LastThird,Acquisition_CSPlus_LastThird,
         Extinction_CSMinus_LastThird,Extinction_CSPlus_LastThird) %>%
  rename(AUD=AUD4orMore) %>%
  pivot_longer(cols = !c(ID,AUD)) %>%
  separate(name, c("Phase","Stimulus","Variable"),convert = T) %>%
  filter(complete.cases(.)) %>%
  group_by(AUD, Phase, Stimulus, Variable) %>%
  summarise(Mean = mean(value,na.rm = TRUE),
            SD = sd(value, na.rm = TRUE),
            N = n()) %>%
  mutate(SE = SD/sqrt(N),
         ci_low = Mean - qt(1-(0.05/2),N-1) * SE,
         ci_high = Mean + qt(1-(0.05/2),N-1) * SE) 


expectancy_plot_data$AUD <- factor(expectancy_plot_data$AUD,
                                                    labels = c("Trauma-exposed control","AUD"))

names(expectancy_plot_data)[1] <- "case_status"


expectancy_plot_data 
```
### bar plots: whole phase mean

```{r}
AUD_bar <-  ggplot(expectancy_plot_data %>%
                          filter(Variable == "Mean"),
                     aes(case_status, Mean,
                         fill = Stimulus))  +
  geom_bar(position = "dodge",
           width = 0.9,
           stat="identity") +
  geom_errorbar(aes(ymin =  Mean - SE,
                    ymax = Mean + SE),
                position = position_dodge(width=.9),
                width = .1)                          +
  facet_wrap(. ~ Phase,
             nrow =2)                              +
  theme_bw()                                         +
  scale_fill_manual(values=PMpallette)                +        
  labs(title=paste("Mean expectancy rating across the whole phase"),
       
         y="Mean expectancy rating\n",
       
       
       subtitle = paste("Bars indicate standard error of the mean\n\ntrauma-exposed controls n = ", 
                        max(expectancy_plot_data$N[expectancy_plot_data$case_status == "Trauma-exposed control"]),
                        "\nAUD n = ",max(expectancy_plot_data$N[expectancy_plot_data$case_status == "AUD"])
                        
                        ))                             +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.key = element_rect(size = 3),
        legend.key.size = unit(3, 'lines'),
        axis.title.y =  element_text(face = "bold",
                                     size = y.title),
        axis.title.x =  element_blank(),
        panel.spacing.x=unit(0, "lines"))     + 
  scale_y_continuous(limits = c(0,9),
                     breaks = c(seq(0,9)))    

AUD_bar

```

### bar plots: first third

```{r}
AUD_bar_first <-  ggplot(expectancy_plot_data %>%
                          filter(Variable == "FirstThird"),
                     aes(case_status, Mean,
                         fill = Stimulus))  +
  geom_bar(position = "dodge",
           width = 0.9,
           stat="identity") +
  geom_errorbar(aes(ymin =  Mean - SE,
                    ymax = Mean + SE),
                position = position_dodge(width=.9),
                width = .1)                          +
  facet_wrap(. ~ Phase,
             nrow =2)                              +
  theme_bw()                                         +
  scale_fill_manual(values=PMpallette)                +        
  labs(title=paste("Mean expectancy rating across the First third of the phase"),
       
         y="Mean expectancy rating\n",
       
       
       subtitle = paste("Bars indicate standard error of the mean\n\ntrauma-exposed controls n = ", 
                        max(expectancy_plot_data$N[expectancy_plot_data$case_status == "Trauma-exposed control"]),
                        "\nAUD n = ",max(expectancy_plot_data$N[expectancy_plot_data$case_status == "AUD"])
                        
                        ))                             +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.key = element_rect(size = 3),
        legend.key.size = unit(3, 'lines'),
        axis.title.y =  element_text(face = "bold",
                                     size = y.title),
        axis.title.x =  element_blank(),
        panel.spacing.x=unit(0, "lines"))     + 
  scale_y_continuous(limits = c(0,9),
                     breaks = c(seq(0,9)))    

AUD_bar_first

```
### bar plots: final third

```{r}
AUD_bar_final <-  ggplot(expectancy_plot_data %>%
                          filter(Variable == "LastThird"),
                     aes(case_status, Mean,
                         fill = Stimulus))  +
  geom_bar(position = "dodge",
           width = 0.9,
           stat="identity") +
  geom_errorbar(aes(ymin =  Mean - SE,
                    ymax = Mean + SE),
                position = position_dodge(width=.9),
                width = .1)                          +
  facet_wrap(. ~ Phase,
             nrow =2)                              +
  theme_bw()                                         +
  scale_fill_manual(values=PMpallette)                +        
  labs(title=paste("Mean expectancy rating across the Last third of the phase"),
       
         y="Mean expectancy rating\n",
       
       
       subtitle = paste("Bars indicate standard error of the mean\n\ntrauma-exposed controls n = ", 
                        max(expectancy_plot_data$N[expectancy_plot_data$case_status == "Trauma-exposed control"]),
                        "\nAUD n = ",max(expectancy_plot_data$N[expectancy_plot_data$case_status == "AUD"])
                        ))                             +
  theme(plot.subtitle = element_text(face = "italic",
                                     size = plot.subtitle),
        plot.title = element_text(face = "italic",
                                     size = plot.title),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(face = "bold",
                                 size = axis.text),
        strip.text = element_text(face = "bold",
                                  size = strip.text),
        legend.key = element_rect(size = 3),
        legend.key.size = unit(3, 'lines'),
        axis.title.y =  element_text(face = "bold",
                                     size = y.title),
        axis.title.x =  element_blank(),
        panel.spacing.x=unit(0, "lines"))     + 
  scale_y_continuous(limits = c(0,9),
                     breaks = c(seq(0,9)))    

AUD_bar_final

```

# save plots

```{r}

# save  figures in folder

## New Group 4
png(paste0(figure_path,"/BarPlot_WholePhaseMeans_NewGroup4.png"),width=18,height=10,unit=units,res=res)
ggdraw(newGroup_bar)
dev.off()

png(paste0(figure_path,"/BarPlot_FirstThirdMeans_NewGroup4.png"),width=18,height=10,unit=units,res=res)
ggdraw(newGroup_bar_first)
dev.off()

png(paste0(figure_path,"/BarPlot_LastThirdMeans_NewGroup4.png"),width=18,height=10,unit=units,res=res)
ggdraw(newGroup_bar_final)
dev.off()

## PTSD

png(paste0(figure_path,"/BarPlot_WholePhaseMeans_PTSD.png"),width=18,height=10,unit=units,res=res)
ggdraw(PTSD_bar)
dev.off()

png(paste0(figure_path,"/BarPlot_FirstThirdMeans_PTSD.png"),width=18,height=10,unit=units,res=res)
ggdraw(PTSD_bar_first)
dev.off()

png(paste0(figure_path,"/BarPlot_LastThirdMeans_PTSD.png"),width=18,height=10,unit=units,res=res)
ggdraw(PTSD_bar_final)
dev.off()

## AUD
png(paste0(figure_path,"/BarPlot_WholePhaseMeans_AUD.png"),width=18,height=10,unit=units,res=res)
ggdraw(AUD_bar)
dev.off()

png(paste0(figure_path,"/BarPlot_FirstThirdMeans_AUD.png"),width=18,height=10,unit=units,res=res)
ggdraw(AUD_bar_first)
dev.off()

png(paste0(figure_path,"/BarPlot_LastThirdMeans_AUD.png"),width=18,height=10,unit=units,res=res)
ggdraw(AUD_bar_final)
dev.off()



```

