---
title: "Step two: impute missed trial data, create summary scores and merge exclusion data with fear conditioning data"
author: "K L Purves"
date: '`r format(Sys.time(), "%d %B, %Y")`'

output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: cerulean
code_folding: show

html_notebook:
  theme: cerulean
toc: yes
---

# Description

This script will read in the flare app data from Katie Bountress with additional grouping variables and will then impute missing trial data with the following logic:

1. If the first trial is missed, the average for the first trial for the same stimulus type (CS+ or CS-) for the whole sample will be used as the missing value   
2. If the last trial is missed, the preceding value for the same stimulus type will be carried forward to replace the missing value 
3. If any other trial is missed, the average of the preceding and following value will be used

Next, summary scores will be created        

1. Means per stimulus per phase
2. Means per stimulus for the first and last third of each phase    
3. Average differential per phase (CS+ - CS- average)    
4. Average differential for first and final third of each phase (CS+ - CS- average for each third)    


Exclusion data will be merged onto the new imputed data file for use in the next steps.

# Setup
This section will set up your environment ready to begin processing the data.


#### Clear global environment
```{r Clear global environment}
remove(list = ls())
```

#### load libraries 
This code block will check if you have the appropriate libraries installed. If you do not, it will download and install them for you. If you do, it will simply call them into your local library.

```{r Setup, libraries}

if(!require(zoo)){
  install.packages("zoo")
  library(zoo)
}

if(!require(plyr)){
  install.packages("plyr")
  library(plyr)
}

if(!require(dplyr)){
  install.packages("dplyr")
  library(dplyr)
}

if(!require(tidyr)){
  install.packages("tidyr")
  library(tidyr)
}


```

#### load in new data with group allocations as csv
converted to csv in spss earlier

This block will ask you to find the *fear conditioning data with correct group allocation as a csv file* on your computer. it will use this information to identify the right path for the finished stage one data set

```{r get files and paths}

## file picker
filename_one <- file.choose()

## get the path to the directory where we will save the finished dataset
save_path <- file.path(dirname(filename_one),"Final_datasets")

## exclusion path
exclusion_path <- file.path(dirname(filename_one),"Exclusions")
  
```

#### read in stage zero data and exclusions dataset


```{r read in data}

data <- read.csv(filename_one)

exclusion_data <- read.csv(paste0(exclusion_path,"/possible_exclusion_dataset.csv"))

```

# Impute missing trial data

## middle trial imputation function

this function will be used later for mid trial imputations. Included here as annoation indicates rules and approaches.
```{r function to impute mid values}
# function to impute missing data from middle trials
# takes as input: index position of the 2nd and second from last stimulus for imputation

imputeMiddle <- function(data,firstPos,lastPos){
  
for (row in 1:dim(data)[1]){
  for (n in firstPos:lastPos){
    
    
    # for the second stimulus, only use the preceding value. If it is NA, then use an average of all the 1st values as the preceding value.
    if(n == firstPos){

      pre <- ifelse(is.na(data[row,n-1]),colMeans(data[n-1]),data[row,n-1])   # If the column before is NA, use the average of the first column across the whole sample
      post <- ifelse(is.na(data[row,n+1]),data[row,n+2],data[row,n+1])  # If the column after is NA, use the one two after.      
      newval <- ifelse(is.na(data[row,n]),(pre + post)/2,data[row,n])
      
      data[row,n] <- newval

    }
    # if the last column is NA, carry forward the last available value
    else if(n == lastPos){

      
      pre <- ifelse(is.na(data[row,n-1]),data[row,n-2],data[row,n-1])   # If the column before is NA, use the average of the first column across the whole sample
      post <- ifelse(is.na(data[row,n+1]),pre,data[row,n+1])  # If the column after is NA, carry forward the last available data point     
      newval <- ifelse(is.na(data[row,n]),(pre + post)/2,data[row,n])

      
      data[row,n] <- newval
    }
    

    else if(is.na(data[row,n])){

      pre <- ifelse(is.na(data[row,n-1]),data[row,n-2],data[row,n-1])   # If the column before is NA, use the one two before.
      post <- ifelse(is.na(data[row,n+1]),data[row,n+2],data[row,n+1])  # If the column after is NA, use the one two after.      
      newval <- ifelse(is.na(data[row,n]),(pre + post)/2,data[row,n])
    

      data[row,n] <- newval

    }
  }
}
return(data)
}

```


## Last trial imputation function
na.locf doesnt work because we are doing row-wise, not column wise. Easier to use function I wrote for pervious data than make data long form and then wide again

Also allows easy specification of column subsets

```{r carry forward last observation function}

imputeLast <- function(dat,lastColIndex) {
  
  for (row in 1:dim(dat)[1]){
    
    if(is.na(dat[row,lastColIndex])){
      
      lastval <- dat[row,lastColIndex-1]
      
      dat[row,lastColIndex] <- lastval
    }
    
    else {
      
      dat[row,lastColIndex] <- dat[row,lastColIndex]
      
    }
  }
  return(dat)
}

```

## Identify any people who missed more than 50% of trials for any stimulus type

And add this a variable for later exclusion in our dataset. Create seperate variables for 
1. number of trials missed during the task (Missing.didNotRespond)
2. number of trials missed due to drop out (Missing.dropOut)
3. Total number of trials missed (Missing.Total)


*Acquisition*
```{r id missing 50%}

acq_minus_missing <- data %>%
  select(ID,
         contains("Acquisition_CSminus")) %>%
  mutate(Acquisition_CSm_Missing.didNotRespond = rowSums(. == 999,na.rm = T),
         Acquisition_CSm_Missing.dropOut = rowSums(is.na(.)),
         Acquisition_CSm_Missing.Total = Acquisition_CSm_Missing.didNotRespond + Acquisition_CSm_Missing.dropOut) %>%
  select(ID,
         Acquisition_CSm_Missing.didNotRespond,
         Acquisition_CSm_Missing.dropOut,
         Acquisition_CSm_Missing.Total)


acq_plus_missing <- data %>%
  select(ID,
         contains("Acquisition_CSplus")) %>%
  mutate(Acquisition_CSp_Missing.didNotRespond = rowSums(. == 999,na.rm = T),
         Acquisition_CSp_Missing.dropOut = rowSums(is.na(.)),
         Acquisition_CSp_Missing.Total = Acquisition_CSp_Missing.didNotRespond + Acquisition_CSp_Missing.dropOut) %>%
  select(ID,
         Acquisition_CSp_Missing.didNotRespond,
         Acquisition_CSp_Missing.dropOut,
         Acquisition_CSp_Missing.Total)


acq_sets <- list(data,
              acq_plus_missing,
              acq_minus_missing)

data <- join_all(acq_sets)

```

*Extinction* 
```{r id missing 50% ext}

ext_minus_missing <- data %>%
  select(ID,
         contains("Extinction_CSminus")) %>%
  mutate(Extinction_CSm_Missing.didNotRespond = rowSums(. == 999,na.rm = T),
         Extinction_CSm_Missing.dropOut = rowSums(is.na(.)),
         Extinction_CSm_Missing.Total = Extinction_CSm_Missing.didNotRespond + Extinction_CSm_Missing.dropOut) %>%
  select(ID,
         Extinction_CSm_Missing.didNotRespond,
         Extinction_CSm_Missing.dropOut,
         Extinction_CSm_Missing.Total)


ext_plus_missing <- data %>%
  select(ID,
         contains("Extinction_CSplus")) %>%
  mutate(Extinction_CSp_Missing.didNotRespond = rowSums(. == 999,na.rm = T),
         Extinction_CSp_Missing.dropOut = rowSums(is.na(.)),
         Extinction_CSp_Missing.Total = Extinction_CSp_Missing.didNotRespond + Extinction_CSp_Missing.dropOut) %>%
  select(ID,
         Extinction_CSp_Missing.didNotRespond,
         Extinction_CSp_Missing.dropOut,
         Extinction_CSp_Missing.Total)


ext_sets <- list(data,
              ext_plus_missing,
              ext_minus_missing)

data <- join_all(ext_sets)

```

*Return of Fear*
```{r id missing 50% ROF}

rof_minus_missing <- data %>%
  select(ID,
         contains("ReturnOfFear_CSminus")) %>%
  mutate(ReturnOfFear_CSm_Missing.didNotRespond = rowSums(. == 999,na.rm = T),
         ReturnOfFear_CSm_Missing.dropOut = rowSums(is.na(.)),
         ReturnOfFear_CSm_Missing.Total = ReturnOfFear_CSm_Missing.didNotRespond + ReturnOfFear_CSm_Missing.dropOut) %>%
  select(ID,
         ReturnOfFear_CSm_Missing.didNotRespond,
         ReturnOfFear_CSm_Missing.dropOut,
         ReturnOfFear_CSm_Missing.Total)


rof_plus_missing <- data %>%
  select(ID,
         contains("ReturnOfFear_CSplus")) %>%
  mutate(ReturnOfFear_CSp_Missing.didNotRespond = rowSums(. == 999,na.rm = T),
         ReturnOfFear_CSp_Missing.dropOut = rowSums(is.na(.)),
         ReturnOfFear_CSp_Missing.Total = ReturnOfFear_CSp_Missing.didNotRespond + ReturnOfFear_CSp_Missing.dropOut) %>%
  select(ID,
         ReturnOfFear_CSp_Missing.didNotRespond,
         ReturnOfFear_CSp_Missing.dropOut,
         ReturnOfFear_CSp_Missing.Total)


rof_sets <- list(data,
              rof_plus_missing,
              rof_minus_missing)

data <- join_all(rof_sets)

```

## convert 999 to NA for the next steps
Now we have saved counts of how many trials were missed due to drop out vs missed ratings we can convert 999 to NA to avoid any numeric calculation issues with the next steps

```{r convert 999 to NA}

data[data == 999] <- NA_real_

```

## impute first trial per stimulus

Where the first trial has been missed, set the value to be the average first trial value for the same phase and stimulus across the whole sample.

I wont impute this for people who missed more than 50% of trials per stimulus

### Acquistion

Not imputing if missing 6 or more responses for each stimulus (12 total trials per stimulus)
```{r first trial missing set to average of whole sample}

# get the average column value and save this as a variable rounded to the nearest whole number

csm1mean <- data %>%
  select(Acquisition_CSminus_1) %>% 
  drop_na() %>%
  summarise_if(is.numeric, mean)

csm1mean <- round(csm1mean$Acquisition_CSminus_1,0)

csp1mean <- data %>%
  select(Acquisition_CSplus_1) %>% 
  drop_na() %>%
  summarise_if(is.numeric, mean)

csp1mean <- round(csp1mean$Acquisition_CSplus_1,0)

data <- data %>%
  mutate(Acquisition_CSminus_1 = case_when(!is.na(Acquisition_CSminus_1) ~ Acquisition_CSminus_1,
                                           is.na(Acquisition_CSminus_1) & Acquisition_CSm_Missing.Total >5 ~ NA_real_,
                                           is.na(Acquisition_CSminus_1) & Acquisition_CSm_Missing.Total < 6 ~ csm1mean),
         Acquisition_CSplus_1 = case_when(!is.na(Acquisition_CSplus_1) ~ Acquisition_CSplus_1,
                                           is.na(Acquisition_CSplus_1) & Acquisition_CSp_Missing.Total >5 ~ NA_real_,
                                           is.na(Acquisition_CSplus_1) & Acquisition_CSp_Missing.Total < 6 ~ csp1mean)
         )
         

```

### Extinction

Not imputing if missing 9 or more responses for each stimulus (18 total trials per stimulus)
```{r first trial missing set to average of whole sample extinction}

# get the average column value and save this as a variable rounded to the nearest whole number
csm1mean <- data %>%
  select(Extinction_CSminus_1) %>% 
  drop_na() %>%
  summarise_if(is.numeric, mean)

csm1mean <- round(csm1mean$Extinction_CSminus_1,0)

csp1mean <- data %>%
  select(Extinction_CSplus_1) %>% 
  drop_na() %>%
  summarise_if(is.numeric, mean)

csp1mean <- round(csp1mean$Extinction_CSplus_1,0)

#convert from integer to numeric for consistent typing
data$Extinction_CSplus_1 <- as.numeric(data$Extinction_CSplus_1)

# impute first values
data <- data %>%
  mutate(Extinction_CSminus_1 = case_when(!is.na(Extinction_CSminus_1) ~ Extinction_CSminus_1,
                                           is.na(Extinction_CSminus_1) & Extinction_CSm_Missing.Total >8 ~ NA_real_,
                                           is.na(Extinction_CSminus_1) & Extinction_CSm_Missing.Total < 9 ~ csm1mean),
         Extinction_CSplus_1 = case_when(!is.na(Extinction_CSplus_1) ~ Extinction_CSplus_1,
                                           is.na(Extinction_CSplus_1) & Extinction_CSp_Missing.Total >8 ~ NA_real_,
                                           is.na(Extinction_CSplus_1) & Extinction_CSp_Missing.Total < 9 ~ csp1mean))
         

```

### Return of fear

Not imputing if missing 2 or more responses for each stimulus (4 total trials per stimulus)
```{r first trial missing set to average of whole sample ReturnOfFear}

# get the average column value and save this as a variable rounded to the nearest whole number
csm1mean <- data %>%
  select(ReturnOfFear_CSminus_1) %>% 
  drop_na() %>%
  summarise_if(is.numeric, mean)

csm1mean <- round(csm1mean$ReturnOfFear_CSminus_1,0)

csp1mean <- data %>%
  select(ReturnOfFear_CSplus_1) %>% 
  drop_na() %>%
  summarise_if(is.numeric, mean)

csp1mean <- round(csp1mean$ReturnOfFear_CSplus_1,0)

#convert from integer to numeric for consistent typing
data$ReturnOfFear_CSminus_1 <- as.numeric(data$ReturnOfFear_CSminus_1)

# impute first values

data <- data %>%
  mutate(ReturnOfFear_CSminus_1 = case_when(!is.na(ReturnOfFear_CSminus_1) ~ ReturnOfFear_CSminus_1,
                                           is.na(ReturnOfFear_CSminus_1) & ReturnOfFear_CSm_Missing.Total >1 ~ NA_real_,
                                           is.na(ReturnOfFear_CSminus_1) & ReturnOfFear_CSm_Missing.Total < 2 ~ csm1mean),
         ReturnOfFear_CSplus_1 = case_when(!is.na(ReturnOfFear_CSplus_1) ~ ReturnOfFear_CSplus_1,
                                           is.na(ReturnOfFear_CSplus_1) & ReturnOfFear_CSp_Missing.Total >1 ~ NA_real_,
                                           is.na(ReturnOfFear_CSplus_1) & ReturnOfFear_CSp_Missing.Total < 2 ~ csp1mean)
         )
         

```

## impute mid trials

any missing trial will be imputed as the average of the preceding and subsequent value using function defined at the beginning of this section.

If the preceding value was NA, then the value two trials before is used.Same for if the following value was also NA. For the 2nd trials where there is only one possible value preceding, if that value is NA, the preceding value is set as the sample mean for trial 1. For the second to last trial, where it is ony possible to have one remaining value after, if this is NA, then the value is carried forward from the last available trial data. 


Note that this is run on the second up to second till last stimulus

### Acquisition
```{r impute mid trials Acquisition }
##  CS MINUS
# get index position of the first relevant column
firstIndex <- grep("^Acquisition_CSminus_2$", names(data))

# get index position of the last relevant column
lastIndex <- grep("^Acquisition_CSminus_11$", names(data))

# apply our imputation function to relevant stimulus columns
data <- imputeMiddle(data,firstIndex,lastIndex)

##  CS PLUS
# get index position of the first relevant column
firstIndex <- grep("^Acquisition_CSplus_2$", names(data))

# get index position of the last relevant column
lastIndex <- grep("^Acquisition_CSplus_11$", names(data))

# apply our imputation function to relevant stimulus columns
data <- imputeMiddle(data,firstIndex,lastIndex)
```

### Extinction
```{r impute mid trials Extinction }
##  CS MINUS
# get index position of the first relevant column
firstIndex <- grep("^Extinction_CSminus_2$", names(data))

# get index position of the last relevant column
lastIndex <- grep("^Extinction_CSminus_17$", names(data))

# apply our imputation function to relevant stimulus columns
data <- imputeMiddle(data,firstIndex,lastIndex)

##  CS PLUS
# get index position of the first relevant column
firstIndex <- grep("^Extinction_CSplus_2$", names(data))

# get index position of the last relevant column
lastIndex <- grep("^Extinction_CSplus_17$", names(data))

# apply our imputation function to relevant stimulus columns
data <- imputeMiddle(data,firstIndex,lastIndex)
```

### Return of Fear
```{r impute mid trials ReturnOfFear }
##  CS MINUS
# get index position of the first relevant column
firstIndex <- grep("^ReturnOfFear_CSminus_2$", names(data))

# get index position of the last relevant column
lastIndex <- grep("^ReturnOfFear_CSminus_3$", names(data))

# apply our imputation function to relevant stimulus columns
data <- imputeMiddle(data,firstIndex,lastIndex)

##  CS PLUS
# get index position of the first relevant column
firstIndex <- grep("^ReturnOfFear_CSplus_2$", names(data))

# get index position of the last relevant column
lastIndex <- grep("^ReturnOfFear_CSplus_3$", names(data))

# apply our imputation function to relevant stimulus columns
data <- imputeMiddle(data,firstIndex,lastIndex)
```

## Impute last trial 

### Acquisition
```{r impute last trials Acquisition }
##  CS MINUS
# get index position of the last relevant column
lastIndex <- grep("^Acquisition_CSminus_12$", names(data))

# apply our imputation function to relevant stimulus columns
data <- imputeLast(data,lastIndex)

##  CS PLUS
# get index position of the last relevant column
lastIndex <- grep("^Acquisition_CSplus_12$", names(data))

# apply our imputation function to relevant stimulus columns
data <- imputeLast(data,lastIndex)
```

### Extinction
```{r impute last trials Extinction }
##  CS MINUS
# get index position of the last relevant column
lastIndex <- grep("^Extinction_CSminus_18$", names(data))

# apply our imputation function to relevant stimulus columns
data <- imputeLast(data,lastIndex)

##  CS PLUS
# get index position of the last relevant column
lastIndex <- grep("^Extinction_CSplus_18$", names(data))

# apply our imputation function to relevant stimulus columns
data <- imputeLast(data,lastIndex)
```

### ReturnOfFear
```{r impute last trials ReturnOfFear }
##  CS MINUS
# get index position of the last relevant column
lastIndex <- grep("^ReturnOfFear_CSminus_4$", names(data))

# apply our imputation function to relevant stimulus columns
data <- imputeLast(data,lastIndex)

##  CS PLUS
# get index position of the last relevant column
lastIndex <- grep("^ReturnOfFear_CSplus_4$", names(data))

# apply our imputation function to relevant stimulus columns
data <- imputeLast(data,lastIndex)
```


# Creat summary scores

## Whole phase means

The average expectancy rating for each stimulus for each phase per person

```{r whole phase means}


acqCSpColnames <- names(data[grep("^Acquisition_CSplus_", names(data))])
acqCSmColnames <- names(data[grep("^Acquisition_CSminus_", names(data))])

extCSpColnames <- names(data[grep("^Extinction_CSplus_", names(data))])
extCSmColnames <- names(data[grep("^Extinction_CSminus_", names(data))])

rofCSpColnames <- names(data[grep("^ReturnOfFear_CSplus_", names(data))])
rofCSmColnames <- names(data[grep("^ReturnOfFear_CSminus_", names(data))])

data <- data %>%
  mutate(Acquisition_CSPlus_Mean = round(rowMeans(select(., all_of(acqCSpColnames))),2),
         Acquisition_CSMinus_Mean = round(rowMeans(select(., all_of(acqCSmColnames))),2),
         Extinction_CSPlus_Mean = round(rowMeans(select(., all_of(extCSpColnames))),2),
         Extinction_CSMinus_Mean = round(rowMeans(select(., all_of(extCSmColnames))),2),
         ReturnOfFear_CSPlus_Mean = round(rowMeans(select(., all_of(rofCSpColnames))),2),
         ReturnOfFear_CSMinus_Mean = round(rowMeans(select(., all_of(rofCSmColnames))),2))

```

## First third means

The average expectancy rating for each stimulus for the first third of each phase (Acquistion and extinction only, ROF has only 4 trials)

```{r first third phase means}


acqCSpColnames.1st <- acqCSpColnames[1:4]
acqCSmColnames.1st <- acqCSmColnames[1:4]

extCSpColnames.1st <- extCSpColnames[1:6]
extCSmColnames.1st <- extCSmColnames[1:6]


data <- data %>%
  mutate(Acquisition_CSPlus_FirstThird = round(rowMeans(select(., all_of(acqCSpColnames.1st))),2),
         Acquisition_CSMinus_FirstThird = round(rowMeans(select(., all_of(acqCSmColnames.1st))),2),
         Extinction_CSPlus_FirstThird = round(rowMeans(select(., all_of(extCSpColnames.1st))),2),
         Extinction_CSMinus_FirstThird = round(rowMeans(select(., all_of(extCSmColnames.1st))),2))

```

## Final third means

The average expectancy rating for each stimulus for the last third of each phase (Acquistion and extinction only, ROF has only 4 trials)

```{r last third phase means}


acqCSpColnames.last <- acqCSpColnames[9:12]
acqCSmColnames.last <- acqCSmColnames[9:12]

extCSpColnames.last <- extCSpColnames[13:18]
extCSmColnames.last <- extCSmColnames[13:18]


data <- data %>%
  mutate(Acquisition_CSPlus_LastThird = round(rowMeans(select(., all_of(acqCSpColnames.last))),2),
         Acquisition_CSMinus_LastThird = round(rowMeans(select(., all_of(acqCSmColnames.last))),2),
         Extinction_CSPlus_LastThird = round(rowMeans(select(., all_of(extCSpColnames.last))),2),
         Extinction_CSMinus_LastThird = round(rowMeans(select(., all_of(extCSmColnames.last))),2))

```

## Overall means differential

the difference between average CS Plus and CS minus ratings per person per phase

```{r whole phase mean differentials}

data <- data %>%
  mutate(Acquisition_WholePhaseDifferential = Acquisition_CSPlus_Mean - Acquisition_CSMinus_Mean,
         Extinction_WholePhaseDifferential = Extinction_CSPlus_Mean - Extinction_CSMinus_Mean,
         ReturnOfFear_WholePhaseDifferential = ReturnOfFear_CSPlus_Mean - ReturnOfFear_CSMinus_Mean)

```

## First third differential

the difference between average CS plus and CS minus ratings per person for the first third of trials (Acquisition and extinction)

```{r first third mean differentials}

data <- data %>%
  mutate(Acquisition_FirstThirdDifferential = Acquisition_CSPlus_FirstThird - Acquisition_CSMinus_FirstThird,
         Extinction_FirstThirdDifferential = Extinction_CSPlus_FirstThird - Extinction_CSMinus_FirstThird)

```

## Last third differential

the difference between average CS plus and CS minus ratings per person for the first third of trials (Acquisition and extinction)

```{r final third mean differentials}

data <- data %>%
  mutate(Acquisition_LastThirdDifferential = Acquisition_CSPlus_LastThird - Acquisition_CSMinus_LastThird,
         Extinction_LastThirdDifferential = Extinction_CSPlus_LastThird - Extinction_CSMinus_LastThird)

```

# save new dataset

```{r save new exclusions}

write.csv(data,paste0(save_path,"/FLAReApp_WithIDS_and_sumscores.csv"),row.names = FALSE)


```

#### Clear global environment
```{r Clear global environment}
remove(list = ls())
```

